# Management Procedures {#MPs}
The purpose of an MSE is to compare the performance of alternative management approaches, or ***Management Procedures*** to identify the method that is most likely to meet the management objectives for the fishery.

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(DLMtool) # only needed for building individual chapter of user guide
for(i in 1:length(DLMdat)) assign(DLMdat[[i]]@Name, DLMdat[[i]])
```

## What is a Management Procedure?
In essence, a Management Procedure is simply a set of rules which define how a fishery will be managed.  These rules can range from simple harvest policies to more complex arrangements.  

For example, a simple Management Procedure may be a constant catch policy, where the annual total allowable catch (TAC) is set a some fixed value.  Alternatively, a more complex Management Procedure may involve multiple data sources, with rules that increase or reduce the TAC in response to trends in one or several indicators.

Management Procedures can differ in data requirements and complexity.  However, all Management Procedures have one thing in common.  They take fishery information and return a management recommendation.

To be included in an MSE, a Management Procedure must be reproducible and able to be coded in a set of instructions.  While fisheries are sometimes managed by expert judgment, it is difficult to reproduce the subjective decision-making process in a computer simulation and include such methods in an MSE.

## Management Procedure Classes
Management Procedures in the DLMtool are divided into two classes: **Output controls** and **Input controls**.

### Output Controls
The output control methods in the DLMtool provide a management recommendation in the form of a TAC.  Some output controls are stochastic, allowing for uncertainty in the data or input parameters, and return a distribution of recommended TACs.

Output control methods are very common in fisheries management, especially in regions which have a tradition of managing fisheries by regulating the total amount of catch.

### Input Controls
The input control methods in the DLMtool allow regulation of **fishing effort**, **size selecitivty**, or **spatial area**.

#### Effort Controls
Effort controls adjust the relative amount of fishing effort. In the DLMtool, all effort controls are designed to adjust effort relative to the current levels of effort.  For example, an input control method may recommend a 15% increase or decrease in fishing effort in response to the signal in some fishery data.

#### Selectivity Controls
Selectivity controls provide recommendations on the size selection pattern for the fishery.  Selectivity controls can be static, for example a single fixed size limit, or dynamic, where a selectivity pattern is modified in response to data.

The DLMtool can account for selectivity controls in the form of a minimum legal length or a harvest slot limit (upper size limit).  A fixed size limit is a Management Procedure, but typically does not require any fishery data to implement the rule.

#### Spatial Controls 
Spatial controls in the DLMtool allow for the opening and closing of spatial areas to fishing activity.  Spatial controls can be static (permanent closed area) or dynamic with an area opening and closing to fishing in response to indicators in the fishery data.

## Built-in Management Procedures
The DLMtool has a large number of Management Procedures that are built into the Toolkit.  Many of these methods have been published in the scientific literature as recommended methods for managing data-limited fisheries.  Other methods have been developed more recently and have been included in the DLMtool to evaluate if they have potential to be useful for management.

The DLMtool has been designed to be extensible, and it is straightforward to develop additional Management Procedures and include them in the MSE.

Details on how to develop new Management Procedures are provided in sections below.

### Output Controls
Output control methods (methods that return a TAC) in the DLMtool are of the class `DLM_output`.  You can use the `avail` function to list the available output controls in the DLMtool:

```{r}
avail("DLM_output")
```

```{r, include=FALSE}
nOut <- length(avail("DLM_output"))
nIn <- length(avail("DLM_input"))
```

As you can see, there are `r nOut` output control methods built into the DLMtool.  

### Input Controls
The list of the `r nIn` available input control methods (class `DLM_input`) can be displayed in the same way:
```{r}
avail("DLM_input")
```

## Details of the Methods
Information on each Management Procedure can be found be using the `?` function. For example, to access the documentation associated with the `CC1` method, you would type `?CC1` into the R console. 

The help documentation provides a description of the method, the arguments for the function, and the output of the function, together with references to the scientific literature that are associated with the method.

## Designing New Methods 
DLMtool was designed to be extensible in order to promote the development of new Management Procedures. In this section we design a series of new Management Procedures that include spatial controls and input controls in the form of size limit restrictions. The central requirement of any MP is that it can be applied to a `DLM_data` object using the function `sapply` (`sfSapply()` in parallel processing).

`DLM_data` objects have a single position `x` for each data entry (e.g. one value for natural mortality rate, a single vector of historical catches, etc.). In the MSE analysis this is extended to `nsim` positions. 

See the [Managing Fishery Data] section for more information about the `DLM_data` object.

It follows that any MP arranged to function `sapply(x,MP,DLM_data)` will work. For example we can get five stochastic samples of the TAC for the demographic FMSY MP paired to catch-curve analysis `FdemCC` applied to a real data-limited data object for red snapper using:
```{r}
sapply(1,Fdem_CC,Red_snapper,reps=5)
```

The MSE populates a `DLM_data` object with many simulations and uses `sfSapply()` (snowfall cluster computing equivalent) to calculate a management recommendation for each simulation. By making methods compatible with this standard the very same equations are used in both the MSE and the real management advice. 

### Average Historical Catch Management Procedure
The average historical catch has been suggested as a starting point for setting TACs in the most data-limited situations (following Restrepo et al., 1998). 

Here we design such an Management Procedure:
```{r}
AvC <- function(x, DLM_data, reps) {
  rlnorm(reps, log(mean(DLM_data@Cat[x,], na.rm=T)), 0.1) 
}
```

Note that all Management Procedures have to be stochastic in this framework, which is why we sample from a log-normal distribution with a CV of roughly 10 per cent. 

Before the Management Procedure can be 'seen' by the rest of the DLMtool package we have to do three more things. 

The  Management Procedure must be assigned a class based on what outputs it provides. Since this is an output control (TAC) based MP we assign it class `DLM_output`. The Management Procedure must also be assigned to the DLMtool namespace:
```{r}
class(AvC) <-"DLM_output"
environment(AvC) <-asNamespace('DLMtool')
```
and - if we are using parallel computing - exported to the cluster:
```{r, eval=FALSE}
sfExport("AvC")
```

<!--
### Third-highest catch
In some data-limited settings third highest historical catch has been suggested as a possible catch-limit. Here we use a similar approach to the average catch MP above (`AvC`) and take draws from a log-normal distribution with CV of 10 per cent:

```{r}
THC <- function(x,DLM_data, reps){
  rlnorm(reps,log(DLM_data@Cat[x,order(DLM_data@Cat[x,],decreasing=T)[3]]),0.1)
}
class(THC)<-"DLM_output"
environment(THC) <- asNamespace('DLMtool')

```
and again export to cluster (if we are using parallel computing):
```{r, eval=FALSE}
sfExport("THC")
```
-->

### Length-at-Selection Set Equal to Length-at-Maturity
To simulate input controls that aim to alter the length-vulnerability to fishing it is possible to design a Management Procedure of class `DLM_input`. 

In this example we set selectivity equal to the maturity curve, while the spatial and effort regulations remain constant:

```{r}
matlenlim <- function (x, DLM_data, ...) {
    Allocate <- 1
    Effort <- 1
    Spatial <- c(1, 1)
    newLFC <- DLM_data@L50[x] * 0.95
    newLFS <- DLM_data@L50[x]
    Vuln <- c(newLFC, newLFS)
    c(Allocate, Effort, Spatial, Vuln)
}
class(matlenlim) <- "DLM_input"
environment(matlenlim) <- asNamespace("DLMtool")

```
and export to cluster:
```{r, eval=FALSE}
sfExport("matlenlim")
```

### Harvest Slot Limit 
We can also choose to create a Management Procedure with a harvest slot limit.  This example sets a minimum legal length at 1.1 times the size of maturity, and a maximum size limit at 75% of the distance between the minimum legal length and the estimate of asymptotic length recorded in the fishery data object:
```{r}
slotlim <- function (x, DLM_data, ...) {
    Allocate <- 1
    Effort <- 1
    Spatial <- c(1, 1)
    newLFS <- 1.1 * DLM_data@L50[x]
    newLFC <- 0.95 * newLFS
    UppLim <- as.numeric(quantile(c(newLFS, DLM_data@vbLinf[x]), 0.75))
    Vuln <- c(newLFC, newLFS, UppLim)
    c(Allocate, Effort, Spatial, Vuln)
}
class(slotlim) <- "DLM_input"
environment(slotlim) <- asNamespace("DLMtool")
```

and export to cluster:
```{r, eval=FALSE}
sfExport("slotlim")
```

Note that the arguments for the input methods must include either `reps` or `...`, even if these are not used. 

### Reducing Fishing Rate in area 1 by 50%
Spatial controls operate similarly to the age/size based controls: a vector of length 2 (the spatial simulator is a 2-box model) that indicates the fraction of current spatial catches. In this example we reduce effort in area 1 by 50%. 

```{r}
area1_50 <- function(x,DLM_data, ...){ 
  Allocate <- 0 # Fraction of effort reallocated to open area
  Effort <- 1  # Fraction of effort in last historical year
  Spatial <- c(0.5,1) # Fraction of effort found in each area
  Vuln <- rep(NA,2) # Length vulnerability is not specified   
  c(Allocate, Effort, Spatial, Vuln) # Input controls stitched togther
}
class(area1_50)<-"DLM_input"
environment(area1_50) <- asNamespace('DLMtool')
```

```{r, eval=FALSE}
sfExport("area1_50")
```