# Custom Parameters (cpars)
By default, DLMtool samples the operating model parameters from a uniform distribution. Because the parameters are sampled independently, it is not possible to generate correlated samples. However, the `cpars` slot in the `OM` object can be used to pass correlated samples into the MSE.

The addition of the `cpars` slot provides a lot of flexibility to the DLMtool, and allows users full control of all parameters used in the model. For example, it is possible to generate operating models directly from the output of common stock assessment packages using functions in DLMtool (e.g `SS2DLM` for Stock Synthesis 3, and `iSCAM2DLM` for a iSCAM model). These functions take the correlated parameter values from the output of the stock assessment and provide them to DLMtool via the `cpars` slot, resulting in an operating model that is conditioned on the stock assessment. 

The `cpars` feature is being continually developed as more features are requested for DLMtool. 

## Valid cpars names
The `cpars` slot requires a named list containing the custom parameter values. You can see the valid names for `cpars` by typing:
```{r}
validcpars()
```

A warning message will alert you if variables appear in the named `cpars` list that are not in `validcpars()`, and these will be ignored in the MSE.

## Correlated samples
As the `cpars` function is used to provide correlated samples to the MSE, it is important that the same number of custom parameters are provided for each variable. In most cases, this is simply a vector `nsim` long. 

For example, if you wish to supply correlated samples of the von Bertalanffy growth parameters, you would create three vectors of length `nsim` containing the samples of Linf, K, and t0.

If the vectors are shorter than `nsim` they will simply be recycled. An error message will alert you if the vectors are not the same length. 

As a demonstration, we will use the `ForceCor` function to generate correlated samples of *M*, *K*, *L50*, and *Linf* and examine the `cpars` slot in the resulting OM object:

```{r}
OM <- ForceCor(DLMtool::testOM)
str(OM@cpars)

```

You can see that the `OM@cpars` slot is a list of length `r length(OM@cpars)` and contains named vectors with `r OM@nsim` correlated samples of the four parameters.  

Because the `OM@cpars` slot contains these values, the *M*, *K*, *L50*, and *Linf* values in the OM, e.g. `OM@M` will be ignored. 

Any additional custom parameters can be added to `cpars` using this same approach. For example, to provide custom (in this case uncorrelated) samples of *t0*:

```{r}
OM@cpars$t0 <- runif(OM@cpars, -1, 0)
str(OM@cpars)
```

## Custom time-varying parameters
It is also possible to supply custom generated time-varing values for some parameters using the `cpars` slot. For example, time-varying natural mortality or selectivity patterns.

Additional details on using the `cpars` slot for this will be added to the userguide soon. If you find that this is a feature you wish to use but are unclear how to do it bug us with an email!

