# Size-Specific Natural Mortality



## Constant M at age/size 
By default DLMtool assumes that natural mortality (*M*) is constant across age and size classes. However, in many species *M* is known to vary by size, and is often assumed to be higher for smaller age-classes and reduces as individuals age and grow. 

A number of users requested the option to include age or size-specific *M* and this has now been added to DLMtool.

There are a number of ways to specify age or size-specific *M* in DLMtool.


## Lorenzen function of weight
Add reference

```{r}
Albacore@Mexp <- c(-0.315, -0.261)

plotM(Albacore)

```

M is scaled so that M at age of maturity is equal to the value sampled from `Stock@M`.

## Map Age-Specifc M


```{r, eval=FALSE}
OM <- new("OM", Blue_shark, Generic_FlatE, Generic_obs, Perfect_Imp)
OM <- ChooseM(OM)
```

```{r, include=FALSE}
OM <- new("OM", Blue_shark, Generic_FlatE, Generic_obs, Perfect_Imp)
OM@M <- c(0.7, 0.65, 0.60, 0.55, 0.616666, 0.6833333, 0.75, 0.63, 0.51, 0.39, .27, 0.15, 0.15, 0.15)
OM@M2 <- c(0.85, 0.8, 0.75, 0.7, 0.766666, 0.8333, 0.9, 0.780, 0.66, 0.54, 0.42, 0.3, 0.3, 0.3, 0.3)
```

```{r}
plotM(OM)
```


## Map Length-Specific M

```{r, eval=FALSE}
OM <- new("OM", Blue_shark, Generic_FlatE, Generic_obs, Perfect_Imp)
OM <- ChooseM(OM, "length")
```

```{r, include=FALSE}
OM <- new("OM", Blue_shark, Generic_FlatE, Generic_obs, Perfect_Imp)
M_at_Length <- read.csv("CSVs/OMcparslength.csv")
OM@cpars$M_at_Length <- M_at_Length

```

```{r}
plotM(OM)
```


