# Selection, Retention and Discard Mortality


## Fishery Selection Curve
The fishery selection or vulnerability to the fishing gear in DLMtool is modelled using a double-normal curve and the parameters in the `Fleet` object: `L5` - smallest length at 5% selection, `LFS` - smallest length at full selection, and `Vmaxlen` the vulnerability of the largest length class (defined as expected length at maximum age `Stock@maxage`).

Here we set up a Operating Model with dome-shaped selectivity and plot a sample of the selectivity-at-age and -length using the `plotSelect` function:

```{r}
OM <- new("OM", Albacore, FlatE_Dom, Generic_obs, Perfect_Imp, nsim=5)
plotSelect(OM, sim=1)
```

The plot shows three curves - vulnerability, realized selection and retention - in each panel. In this case they are all the same, because the default setting of DLMtool is to assume that all selected fish are retained in the catch. 

## Fishery Retention Curve
In some cases the fishing gear selects fish (often small sizes) that are not retained in the catch and are discarded at sea. The fishery-retention curve can be specified following the same approach as selectivity, using the following slots in the `Fleet` or `OM` object:

- `LR5` - the smallest length at 5% retention
- `LFR` - the smallest length at full selection
- `Rmaxlen` - the retention of the largest size class (defined as expected length at maximum age `Stock@maxage`).  

The default values for these parameters are:
```{r}
OM@LR5
OM@LFR
OM@Rmaxlen
```
meaning that the default assumption is that all size classes are fully retained by the fishery. 

The retention curve can be modified by providing values for these slots:

```{r}
OM@LR5 <- c(0.6, 0.7)
OM@LFR <- c(0.9, 1)
```

Note that the values in the `LR5` and `LFR` slots must be in the same units as those in the `L5` and `LFS` slots. Here we are specifying the values relative to the size of maturity, and assuming that the fishery discards the smaller sized fish:
```{r}
plotSelect(OM, sim=1)
```

The plot shows that the retention curve for the fishery has shifted to the right, towards larger and older fish, while the vulnerability of the fishing gear remains the same.

Because we are assuming no discard mortality in this case, the realized selection and retention curves are equivalent. This means that although fish of age/length between the vulnerability and retention curves are selected by the fishery, they are discarded with 100% survival and therefore are not removed from the population.

## Discard Mortality 
The assumption of 100% survival of discarded fish may be unrealistic in many situations. Discard mortality can be specified by the `Fdisc` slot in the `Stock` or `OM` object. The `Fdisc` slot represents the fraction of discarded fish that die, or $1-\text{survival}$. Here we assume that between 30 and 50% of discarded fish suffer fishing mortality:

```{r}
OM@Fdisc <- c(0.3, 0.5)
plotSelect(OM, sim=1)
```

We can see now that the realized selection and the retention curves are different for the age/size classes that are discarded by the fishery. The realized selection curve (dashed red line) represents the actual selectivity of the fish removed from the population. 

The retention curve (dotted blue line) shows the age/size classes that are retained by the fishery and appear in the total cathc, catch-at-age, and catch-at-length fishery data. 

The shaded gray area between these two curves represents that age/size classes that are caught and killed by the fishery but are discarded and do not appear in the catch statistics. 

The gear vulnerability curve remains unchanged, and shows that some individuals in the smaller age/size classes are caught and discarded alive back into the population.

## General Discarding
General discarding across all age or size classes can be included using the discarding rate slot `DR` in the `Fleet` or `OM` object. 

For example, here we assume that between 10 and 20% of all age/size classes are discarded by the fishery:

```{r}
OM@DR <- c(0.1, 0.2)
```
Plotting the selectivity and retention curves shows that a proportion of all age and size classes are now discarded, with the survival rate determined by the `Fdisc` parameter:

```{r}
plotSelect(OM, sim=1)
```




